`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks

### 2020-02-14 analysis/paper/sim_survey_paper.Rmd
Tweak abstract and introduction

Fish populations often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective surveys. Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics). For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of fish captured to measure biological characteristics (e.g. lengths or ages). Population estimates are also affected by the pathway taken to analyze such data. Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for fisheries-independent survey data. Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations. The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of built-in or user-defined sampling protocols to reproduce the population parameters of the known population. **`SimSurvey`** also includes a function for estimating the stratified mean and variance of the population from the simulated survey data. We demonstrate the use of this package using a case study and show that the it can reveal unexpected sources of bias and be used to explore design-based solutions to such problems. In summary, **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring. Various statistical approaches can then be applied to the results to test the efficacy of different analytical approaches.

@@

One of the biggest challenges associated with surveys is the costs of planning and implementing effective and efficient sampling designs [@reynolds2011]. Despite the costs, surveys have become a mainstay in the management of dynamic fish stocks as survey data are not easily replaced by other sources of information [@Pennington1998]. For instance, large population declines can be overlooked by relying solely on catch and effort data from commercial fisheries [@rose1999]. Such difficulties have contributed to the growing reliance on fisheries-independent surveys to monitor fish populations. Planning or improving such surveys is an important but non-trivial task because of the inherent complexity of both the survey and target population. Specifically, fisheries-independent surveys usually involve multiple levels of sampling and fish populations often show size-structured spatial and temporal dynamics [@Nelson2014]. These features are not easily captured by simple statistical models, which makes it difficult to determine optimal sampling and sub-sampling schemes [@Pennington1994; @Pennington2002; @Stewart2014]. In such situations it is common for practitioners to resort to simulations as a means of testing the efficacy of various sampling designs (e.g. [@thomas2010]), however, the complexities associated with fisheries-independent surveys may explain why simulations that test the full sampling and analytical pathway are rare (but see [@puerta2019; @schnute2003]).

Here we document **`SimSurvey`**, an `R` package designed to simplify and facilitate realistic simulations of fisheries-independent trawl surveys. In short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time. The package has two main components: the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated population distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.

This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al. [-@puerta2019], however, efforts were focused on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways. The steps taken to simulate surveys of spatial, age-structure populations are outlined below. Under the **[Methods]** section, we first outline the equations underlying the package (**[Model structure]** section) and then move into a vignette of the package (**[Core functions]** section). While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study. Output from default function calls are therefore relevant to the case study and some of these results are highlighted in the **[Interpretation]** section; a more detailed description and discussion of these results can be found in **[S1 Appendix][S1 Appendix: Case study]**.

>>>>

Fish populations often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective surveys. Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics). For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of fish captured to measure biological characteristics (e.g. lengths or ages). Population estimates are also affected by the pathway taken to analyze such data. Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for fisheries-independent survey data. Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations. The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of built-in or user-defined sampling protocols to reproduce the population parameters of the known population. **`SimSurvey`** also includes a function for estimating the stratified mean and variance of the population from the simulated survey data. We demonstrate the use of this package using a case study and show that it can reveal unexpected sources of bias and be used to explore design-based solutions to such problems. In summary, **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring. Various statistical approaches can then be applied to the results to test the efficacy of different analytical approaches.

@@

One of the biggest challenges associated with surveys is the costs of planning and implementing effective and efficient sampling designs [@reynolds2011]. Despite the costs, surveys have become a mainstay in the management of dynamic fish stocks as survey data are not easily replaced by other sources of information [@Pennington1998]. For instance, large population declines can be overlooked by relying solely on catch and effort data from commercial fisheries [@rose1999]. Such difficulties have contributed to the growing reliance on fisheries-independent surveys to monitor fish populations. Planning or improving such surveys is an important but non-trivial task because of the inherent complexity of both the survey and target population. Specifically, fisheries-independent surveys usually involve multiple levels of sampling and fish populations often show size-structured spatial and temporal dynamics [@Nelson2014]. These features are not easily captured by simple statistical models, which makes it difficult to determine optimal sampling and sub-sampling schemes [@Pennington1994; @Pennington2002; @Stewart2014]. In such situations it is common for practitioners to resort to simulations to test various sampling and analytical pathways (e.g. [@thomas2010]), however, such tests are rare in fisheries research (but see [@schnute2003; @puerta2019]), perhaps because of the complexities associated with fisheries-independent surveys.

Here we document **`SimSurvey`**, an `R` package designed to facilitate realistic simulations of fisheries-independent trawl surveys. In short, the package allows for the simulation of random or stratified-random surveys of an age-structured population that varies in space and time. The package has two main components: the first focuses on mimicking realistic fish stocks by simulating a spatially and age-correlated population distributed across a habitat gradient, and the second component focuses on simulating various surveys of these virtual fish stocks.

This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al. [-@puerta2019], however, efforts were focused on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways. The steps taken to simulate surveys of spatial, age-structure populations are outlined below. We first outline the equations underlying the package (**[Model structure]** section) and then move into a vignette (**[Core functions]** section). The core functions of the package are largely demonstrated using default settings, and these settings are based on a case study (see for **[S1 Appendix][S1 Appendix: Case study]** details). Several of the results from the case study are described and discussed in the **[Interpretation]** section as they highlight the utility of the package. Finally, we provide some guidance on how to modify default settings to suit specific needs (**[Parameterisation]** section) and then move onto discuss the potential uses and future directions of the package (**[Discussion]** section).
